### 2022-11-01 Papers/ICCC-2023/sections/design.tex
corrections

To declare an attribute in C-=-1, the \lstinline{att} keyword is used.
After that, attribute targets should be listed in angled brackets, as in line 0 of Listing \ref{lst:noDiscardCm1}.
Attributes valid targets for attributes includes a number of language elements, such as types, functions, variables and fields.Attribute from Listing \ref{lst:noDiscardCm1} declares two member functions: \lstinline{attach} in line 4 and \lstinline{onCall} in line 6.

@@

The \lstinline{attach} method is called after names of all of the program elements have been gathered, but before the compiler starts to analyze function bodies.
It is common accross all attribute targets and accepts the descriptor of the attached program element (function, field, type, etc.).
Only during call to \lstinline{attach} can the attribute change aspects of the program that affect function overload resolution, for example
whether a function is invokable at run or compile time.

@@

One of the possible solutions to the above mentioned problem is to limit which functions can be invoked at compiletime.
C-=-1 allows code within a compiletime context to invoke procedures only from other packages, declared explicitly as dependencies.
Circular references between packages, as in most other languages, are forbiden.
C-=-1 additionally prohibits modification of dependencies.
Therefore its is impossible for a function to modify a procedure it depends on.

>>>>

To declare an attribute in C-=-1, the \lstinline{att} keyword is used.
After that, attribute targets should be listed in angled brackets, as in line 0 of Listing \ref{lst:noDiscardCm1}.
Attributes valid targets for attributes includes a number of language elements, such as types, functions, variables and fields. Attribute from Listing \ref{lst:noDiscardCm1} declares two member functions: \lstinline{attach} in line 4 and \lstinline{onCall} in line 6.

@@

The \lstinline{attach} method is called after names of all the program elements have been gathered, but before the compiler starts to analyze function bodies.
It is common accross all attribute targets and accepts the descriptor of the attached program element (function, field, type, etc.).
Only during call to \lstinline{attach} can the attribute change aspects of the program that affect function overload resolution, for example
whether a function is invokable at run or compile time.

@@

One of the possible solutions to the above-mentioned problem is to limit which functions can be invoked at compiletime.
C-=-1 allows code within a compiletime context to invoke procedures only from other packages, declared explicitly as dependencies.
Circular references between packages, as in most other languages, are forbiden.
C-=-1 additionally prohibits modification of dependencies.
Therefore, its is impossible for a function to modify a procedure it depends on.

### 2022-11-01 Papers/ICCC-2023/sections/implementation.tex
corrections

The final part of a CTFEF compiler is the Backend Interface.
It is a program, written in the target language, and ran at compile-time that translates the semantic model into the backends assembly language.
Backend Interface implemented for C-=-1 is relativley small and is described in chapter \ref{implementation/backend-interface}.

@@

Figure \ref{fig:interpreter_data_structures} contains a class diagram of most of the types used to represent values within C-=-1.
All of them derive from \lstinline{IRuntimeValue} and are managed via C++ smart pointers.
The interface of the base class allows the value to be converted to a human readable format, serialisation, deserialisation and copying.

@@

User-defined types are represented using \lstinline{ObjectValue}.
The contents of an objects is kept as a \lstinline{string} - \lstinline{IRuntimeValue} dictionary, with field names as keys and \lstinline{uniqe_ptr<IRuntimeValue>} as values.
C-=-1 object is therefore spread out in memory, even if the fields are directly contained within the class, without any indirection.

@@

A major motivation for creating CTFEF was the ability to support languages with compile-time metaprogramming.
This includes reflection and modification of the code being compiled.
User program has to be represented as a complete and modifiable object, using the interpreters data structures.

@@

To aid in generating meaningfull error messages, all elements of the semantic model, have a \lstinline{sourcePointers}.
It is a simple structure, that contains the filename and the line number of the expression or instruction.
This information can be passed to compiler intrinsic functions, such as \lstinline{raiseError}, to generate error messages for the user.
Listing \ref{lst:noDiscardCm1} contains a example use of this functionality.

@@

Certain other intrinsic operations are defined using external dependencies.
C-=-1 memory managment library, in the runtime context, uses a simple interface capable of allocating and deleting a continous buffer.
It is comprised of two functions: \lstinline{unsafe_new} and \lstinline{delete}.
In the standard library, they are explicitly mapped to \lstinline{malloc} and \lstinline{free} functions from the C runtime.

>>>>

The final part of a CTFEF compiler is the Backend Interface.
It is a program, written in the target language, and ran at compile-time that translates the semantic model into the backends' assembly language.
Backend Interface implemented for C-=-1 is relativley small and is described in chapter \ref{implementation/backend-interface}.

@@

Figure \ref{fig:interpreter_data_structures} contains a class diagram of most of the types used to represent values within C-=-1.
All of them derive from \lstinline{IRuntimeValue} and are managed via C++ smart pointers.
The interface of the base class allows the value to be converted to a human-readable format, serialisation, deserialisation and copying.

@@

User-defined types are represented using \lstinline{ObjectValue}.
The contents of an object is kept as a \lstinline{string} - \lstinline{IRuntimeValue} dictionary, with field names as keys and \lstinline{uniqe_ptr<IRuntimeValue>} as values.
C-=-1 object is therefore spread out in memory, even if the fields are directly contained within the class, without any indirection.

@@

A major motivation for creating CTFEF was the ability to support languages with compile-time metaprogramming.
This includes reflection and modification of the code being compiled.
User program has to be represented as a complete and modifiable object, using the interpreters' data structures.

@@

To aid in generating meaningfull error messages, all elements of the semantic model, have a \lstinline{sourcePointers}.
It is a simple structure, that contains the filename and the line number of the expression or instruction.
This information can be passed to compiler intrinsic functions, such as \lstinline{raiseError}, to generate error messages for the user.
Listing \ref{lst:noDiscardCm1} contains an example use of this functionality.

@@

Certain other intrinsic operations are defined using external dependencies.
C-=-1 memory managment library, in the runtime context, uses a simple interface capable of allocating and deleting a continous buffer.
It consists of two functions: \lstinline{unsafe_new} and \lstinline{delete}.
In the standard library, they are explicitly mapped to \lstinline{malloc} and \lstinline{free} functions from the C runtime.

### 2022-11-01 Papers/ICCC-2023/sections/introduction.tex
corrections

Compile-time function execution (CTFE) is an aspect of a programming language, that allows the programmer to execute code at compile-time.
It has been gaining popularity with low level programming langauges such as Rust \cite{rust} or C++ \cite{ISO:cpp98}.
The goals and capabilities of CTFE depend on the language and are discussed in chapter \ref{related-work}.

In this paper a new approach to compiler construction that places Compile Time Function Execution capabilities as the first priority is proposed.
This architecture is called CTFEF: Compile Time Function Execution First.
CTFEF builds the compiler around the interpreter component.
The role of the compiler is to construct the semantic model of the program being compiled and pass it as data to the interpreted Compiler-Intreface module.
Compiler-Interface then tranforms the model into the backend's assembly language to be compiled to executable.
This approach was created during implementation of a compiler for C-=-1, a new language that prioritized static metaprogramming.
The design of that language is described in chapter \ref{language-design}.
Chapter \ref{compiler-design} describes the structure of a CTFEF compiler and how its components interact.

Using CTFEF approach carries with it certain unique implementation considerations.
Data structures used within the compiler and the way that the program is modeled is tightly coupled with the compiled language.
These challenges, as well as how they were solved in C-=-1 compiler are described in chapter \ref{implementation}.

>>>>

Compile-time function execution (CTFE) is an aspect of a programming language, that allows the programmer to execute code at compile-time. 
It has been gaining popularity with low level programming languages such as Rust \cite{rust} or C++ \cite{ISO:cpp98}.
The goals and capabilities of CTFE depend on the language and are discussed in details in chapter \ref{related-work}.

In this paper a new approach to compiler construction that places Compile Time Function Execution capabilities as the first priority is proposed. 
This architecture is called CTFEF: Compile Time Function Execution First. CTFEF builds the compiler around the interpreter component. 
The role of the compiler is to construct the semantic model of the program being compiled and pass it as data to the interpreted Compiler-Interface module. 
Compiler-Interface then transforms the model into the assembly language to be compiled into executable. 
This approach was created during implementation of the compiler for C-=-1, a new language that prioritizes static metaprogramming. 
The design of that language is described in chapter \ref{language-design}.
Related work is briefly shown in chapter \ref{related-work}.
In chapter \ref{compiler-design} the structure of a CTFEF compiler and interactions among its components are presented.

Using CTFEF approach causes many implementation difficulties. Data structures used within the compiler are tightly coupled with the compiled language. 
These problems and how they were solved in C-=-1 compiler are described in chapter \ref{implementation}.

### 2022-11-01 Papers/ICCC-2023/sections/realted-work.tex
a lot of changes in the article
corrections

Many currently used programming languages allow the programmer to execute some code at compile time, including C\# \cite{csharp:source_generators,roslyn}, Rust \cite{rust, klabnik2019rust} and C++ \cite{ISO:cpp20}.

Rust language compiler is the most similar in capability, to what was demonstrated with C-=-1, as a feature of CTFE First.
It allows the user to write code that performs some transformations of the program, and interact with the envoirment during the build process.
The two relevant features are build scripts and macros.

Rust macros, much like the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens rather than text and use syntax similar to regular Rust code.
This mechanism is powerfull, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain much of the information available to the compiler.

Build scripts, on the other hand, are programs that execute prior to compilation of the main package.
They prepare the envoirment for building the program by, for example, compiling its external dependencies.
They do not have any information about the structure of the program being compiled.
Build scripts that require such data, would have to analyze the code by itself, without any help from the compiler.
This makes application such as automatically genertaing bindings for other programming languages significantly more difficult.

>>>>

Many currently used programming languages allow the programmer to execute some code at compile time, for example: C\# \cite{csharp:source_generators,roslyn}, Rust \cite{rust, klabnik2019rust} and C++ \cite{ISO:cpp20}.

Rust language compiler is the most similar in capability, to what was demonstrated with C-=-1, as a feature of CTFEF.
It allows the user to write code that performs some transformations of the program, and interact with the environment during the build process.
The two relevant features are build scripts and macros.
Rust macros, similar to the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens, rather than text, and use syntax similar to regular Rust code.
This mechanism is powerful, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain some information available to the compiler.

Build scripts, on the other hand, are programs that execute before the compilation of the main package.
They prepare the environment for building the program, for example, compile external dependencies.
The structure of the program being compiled, is not available for build scripts.
Build scripts accessing such data, would have to analyze the code by themselves, without any compiler assistance, and that makes automatic generation of bindings for other programming languages very difficult.
